### 2015-02-27 writing/F'14 paper/discussion.tex
Paper update.

This paper first evaluates two recursive models on a series of three increasingly
challenging interpretive tasks involving natural language inference---the 
core relational algebra of natural logic with entailment and
exclusion; recursive propositional logic structures; and statements
involving quantification and negation. We then show that the same models can learn to
perform an entailment task on natural language image captions. The results suggest that TreeRNTNs,
and potentially also TreeRNNs, have the capacity to learn to model these tasks with 
reasonably-sized training sets. These positive results are
promising for the future of learned representation models in the
applied modeling of compositional semantics.

Some questions about the abilities of these models remain open. Even
the TreeRNTN falls short of perfection in the recursion experiment, with
performance falling off steadily as the size of the expressions grows. It
remains to be seen whether these deficiencies can be overcome with
stronger models or learning techniques. In addition, interesting 
analytical questions remain about \ii{how} these models encode
the underlying logics. Neither the underlying
logical theories, nor any straightforward parameter inspection technique provides 
much insight on this point, but we hope that further experiments may reveal 
some structure in the representations.

>>>>

This paper first evaluates two recursive models on a series of three increasingly
challenging interpretive tasks involving natural language inference---the 
core relational algebra of natural logic with entailment and
exclusion; recursive propositional logic structures; and statements
involving quantification and negation. We then show that the same models can learn to
perform an entailment task on natural language. The results suggest that TreeRNTNs,
and potentially also TreeRNNs, have the capacity to learn to model these tasks with 
reasonably-sized training sets. These positive results are
promising for the future of learned representation models in the
applied modeling of compositional semantics.

Some questions about the abilities of these models remain open. Even
the TreeRNTN falls short of perfection in the recursion experiment, with
performance falling off steadily as the size of the expressions grows. It
remains to be seen whether these deficiencies can be overcome with
stronger models or learning techniques. In addition, interesting 
analytical questions remain about \ii{how} these models encode
the underlying logics. Neither the underlying
logical theories, nor any straightforward parameter inspection technique provides 
much insight on this point, but we hope that further experiments may reveal 
structure in the learned parameters or the representations they produce.

### 2015-02-27 writing/F'14 paper/intro.tex
Paper update.

% TODO: Is there any way to squeeze a useful example into the first page?
% TODO: Can we fit a version of the ``find something that does that'' quote in intro somewhere?

% TODO: This para might be a bit repetitive.
Tree-structured recursive neural network models (TreeRNNs; \citealt{goller1996learning}) for sentence meaning
have been successful in an array of sophisticated language tasks,
including sentiment analysis \cite{socher2011semi,irsoydeep},
image description \cite{sochergrounded}, and paraphrase detection
\cite{Socher-etal:2011:Paraphrase}. These results are encouraging for
the ability of these models to learn to produce and use strong 
semantic representations for sentences,
though many remain skeptical of the extent to which they can learn to represent
the necessary complex compositional structures using fixed-size vectors. 
In particular, it remains an open question whether learned models can achieve the
high-fidelity distributed representations of recent algebraic
approaches to vector space modeling \cite{ClarkCoeckeSadrzadeh2011,grefenstette2013towards,Hermann-etal:2013,rocktaschellow},
and whether they can match the performance of grammars based in logical forms
in their ability to model core semantic phenomena like quantification, entailment, and contradiction \cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013}.

@@

The experiments with simulated data provide a convincing demonstration
of the ability of neural networks to learn to encode semantic
inferences in complex natural language sentences from reasonably-sized
training sets. However, we are also interested in the more practical
question of whether they can learn these representations from
naturalistic text. To address this question, we apply our models to
the SICK entailment challenge data in \S\ref{sec:sick}. The small size of this corpus puts
data-hungry NN models like ours at a disadvantage, but we are
nonetheless able to achieve competitive performance on it,
surpassing several submitted models with significant hand-engineered task-specific features.
This suggests that the representational abilities that we observe in the previous sections
are not limited to carefully circumscribed tasks. We conclude that TreeRNTN models
are adequate for typical cases of natural logic inference, and that there is not yet
any clear level of inferential complexity for which other approaches work and NN models fail.

% TODO: Contributions para.

>>>>

Tree-structured recursive neural network models (TreeRNNs; \citealt{goller1996learning}) for sentence meaning
have been successful in an array of sophisticated language tasks,
including sentiment analysis \cite{socher2011semi,irsoydeep},
image description \cite{sochergrounded}, and paraphrase detection
\cite{Socher-etal:2011:Paraphrase}. These results are encouraging for
the ability of these models to learn to produce and use strong 
semantic representations for sentences, but it remains an 
open question whether any such fully learned model can achieve the kind of
high-fidelity distributed representations proposed in recent algebraic
work on vector space modeling 
\cite{ClarkCoeckeSadrzadeh2011,grefenstette2013towards,Hermann-etal:2013,rocktaschellow},
and whether any such model can match the performance of grammars based in logical forms
in their ability to model core semantic phenomena like quantification, entailment, and contradiction \cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013}.

@@

The experiments with simulated data provide a convincing demonstration
of the ability of neural networks to learn to encode semantic
inferences in complex natural language sentences from reasonably-sized
training sets. However, we are also interested in the more practical
question of whether they can learn these representations from
naturalistic text. To address this question, we apply our models to
the SICK entailment challenge data in \S\ref{sec:sick}. The small size of this corpus puts
data-hungry NN models like ours at a disadvantage, but we are
nonetheless able to achieve competitive performance on it,
surpassing several submitted models with significant hand-engineered task-specific features.
As with our simulated data, our concern is primarily with the quality of the learned representations and their ability to encode our target semantic relations. Thus, we are less focused on absolute performance than on patterns of accurate and inacccurate predictions, so even this small data set can provide powerful analytic insights into our core representational issues.
This suggests that the representational abilities that we observe in the previous sections
are not limited to carefully circumscribed tasks. We conclude that TreeRNTN models
are adequate for typical cases of natural logic inference, and that there is not yet
any clear level of inferential complexity for which other approaches work and NN models fail.

### 2015-02-27 writing/F'14 paper/join.tex
Paper update.

\begin{table}[tp]
  \centering  \small
  \setlength{\arraycolsep}{8pt}
  \renewcommand{\arraystretch}{1.1}
  \newcommand{\UNK}{\cdot}  
  $\begin{array}[t]{c@{ \ }|*{7}{c}|}
    %\hline
    \multicolumn{1}{c}{}
             & \nateq     & \natfor     & \natrev     & \natneg    & \natalt     & \natcov     & \multicolumn{1}{c}{\natind} \\
    \cline{2-8}
    \nateq  & \nateq &   \natfor &  \natrev &  \natneg &   \natalt &  \natcov &  \natind \\
    \natfor & \natfor &  \natfor &  \UNK &  \natalt &   \natalt &  \UNK &  \UNK \\
    \natrev & \natrev &  \UNK &  \natrev &  \natcov &   \UNK &  \natcov &  \UNK \\
    \natneg & \natneg &  \natcov &  \natalt &  \nateq &    \natrev &  \natfor &  \natind \\
    \natalt & \natalt &  \UNK &  \natalt &  \natfor &   \UNK &  \natfor &  \UNK \\
    \natcov & \natcov &  \natcov &  \UNK &  \natrev &   \natrev &  \UNK &  \UNK \\
    \natind & \natind & \UNK &  \UNK &  \natind &  \UNK &  \UNK &  \UNK \\
    \cline{2-8}
  \end{array}$
  \caption{In \S\ref{sec:join}, we assess our models' ability to learn to do inference over pairs of relations using the rules represented here. As an example, given that $p_1 \natfor p_2$ and $p_2 \natneg p_3$, the entry in the $\natfor$ row and the $\natneg$ column lets us conclude that $p_1 \natalt p_3$.
These rules are derived from the definitions of the relations in Table~\ref{b-table}.  Cells containing a dot correspond to pairs
of relations for which no valid inference can be drawn.} 
  \label{tab:jointable}
\end{table}

@@

\paragraph{Experiments}
We begin by creating a world model
on which we will base the statements in the train and test sets.
This takes the form of a small boolean structure in which terms denote
sets of entities from a small domain.  Fig.~\ref{lattice-figure}a
depicts a structure of this form with three entities ($a$, $b$, and $c$) and eight terms ($p_{1\text{-}8}$). We then generate a 
relational statement for each pair of terms in the model, as shown in Fig.~\ref{lattice-figure}b. 
We divide these statements evenly into training and test sets, and remove from the
test set statements which cannot be proven from the training
statements.
In each experimental run, we create a model with 80 terms over a domain of 7 elements, yielding a training set of 3200 relations and a test set of 
2960 relations.

@@

\paragraph{Results} 
Table \ref{joinresultstable} shows the results. 
The NTN is able to accurately encode the relations 
between the terms in the geometric relations between their vectors, 
and is able to then use that information to recover relations that 
are not overtly included in the training data. The NN also generalizes fairly well, 
but makes enough errors that it remains an open question whether 
it is capable of learning representations with these properties. 
We cannot rule out the possibility that different optimization techniques or
further hyperparameter tuning could lead an NN model to succeed here.

>>>>

\begin{table}[tp]
  \centering  \small
  \setlength{\arraycolsep}{8pt}
  \renewcommand{\arraystretch}{1.1}
  \newcommand{\UNK}{\cdot}  
  $\begin{array}[t]{c@{ \ }|*{7}{c}|}
    %\hline
    \multicolumn{1}{c}{}
             & \nateq     & \natfor     & \natrev     & \natneg    & \natalt     & \natcov     & \multicolumn{1}{c}{\natind} \\
    \cline{2-8}
    \nateq  & \nateq &   \natfor &  \natrev &  \natneg &   \natalt &  \natcov &  \natind \\
    \natfor & \natfor &  \natfor &  \UNK &  \natalt &   \natalt &  \UNK &  \UNK \\
    \natrev & \natrev &  \UNK &  \natrev &  \natcov &   \UNK &  \natcov &  \UNK \\
    \natneg & \natneg &  \natcov &  \natalt &  \nateq &    \natrev &  \natfor &  \natind \\
    \natalt & \natalt &  \UNK &  \natalt &  \natfor &   \UNK &  \natfor &  \UNK \\
    \natcov & \natcov &  \natcov &  \UNK &  \natrev &   \natrev &  \UNK &  \UNK \\
    \natind & \natind & \UNK &  \UNK &  \natind &  \UNK &  \UNK &  \UNK \\
    \cline{2-8}
  \end{array}$
  \caption{In \S\ref{sec:join}, we assess our models' ability to learn to do inference over pairs of relations using the rules represented here, which are derived from the definitions of the relations in Table~\ref{b-table}.  As an example, given that $p_1 \natfor p_2$ and $p_2 \natneg p_3$, the entry in the $\natfor$ row and the $\natneg$ column lets us conclude that $p_1 \natalt p_3$. Cells containing a dot correspond to situations for which no valid inference can be drawn.} 
  \label{tab:jointable}
\end{table}

@@

\paragraph{Experiments}
We begin by creating a world model
on which we will base the statements in the train and test sets.
This takes the form of a small boolean structure in which terms denote
sets of entities from a small domain.  Fig.~\ref{lattice-figure}a
depicts a structure of this form with three entities ($a$, $b$, and $c$) and eight terms ($p_{1\text{-}8}$). We then generate a 
relational statement for each pair of terms in the model, as shown in Fig.~\ref{lattice-figure}b. 
We divide these statements evenly into train and test sets, and delete the test set
 examples which cannot be proven from the train examples, for which there is not enough information for even an ideal system to choose a correct label.
In each experimental run, we create a model with 80 terms over a domain of 7 elements, yielding a training set of 3200 relations and a test set of 
2960 relations.

@@

\paragraph{Results} 
The resuts (Table \ref{joinresultstable}) show that NTN is able to accurately encode the relations between the terms in the geometric relations between their vectors, 
and is able to then use that information to recover relations that 
are not overtly included in the training data. The NN also generalizes fairly well, 
but makes enough errors that it remains an open question whether 
it is capable of learning representations with these properties. 
We cannot rule out the possibility that different optimization techniques or
further hyperparameter tuning could lead an NN model to succeed here.

### 2015-02-27 writing/F'14 paper/methods.tex
Paper update.

For the sentence embedding portions of the network, we evaluate both TreeRNN models with
 the standard NN layer function \eqref{TreeRNN} and those with the more powerful neural tensor 
 network layer function
\eqref{TreeRNTN} proposed in \newcite{chen2013learning}. The nonlinearity $f(x) = \tanh(x)$ is applied
 elementwise to the output of either layer function.
%
\begin{gather} 
\label{TreeRNN}
\vec{y}_{\textit{TreeRNN}} = f(\mathbf{M} \colvec{2}{\vec{x}^{(l)}}{\vec{x}^{(r)}} + \vec{b}\,) \\
\label{TreeRNTN} 
\vec{y}_{\textit{TreeRNTN}} = \vec{y}_{\textit{TreeRNN}} + f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)})
\end{gather} 
%
Here, $\vec{x}^{(l)}$ and $\vec{x}^{(r)}$ are the column vector
representations for the left and right children of the node, and
$\vec{y}$ is the node's output.  The TreeRNN concatenates them, multiplies
them by an $n \times 2n$ matrix of learned weights, and adds a bias $\vec{b}$. 
The TreeRNTN adds a learned third-order tensor 
$\mathbf{T}$, of dimension $n \times n \times n$, modeling
multiplicative interactions between the child vectors. 
Other tree-structured networks have been successfully used in past work
\cite{sochergrounded,kalchbrenner2014convolutional,irsoydeep}, but
we believe that these two provide a valuable case study, and that positive results on 
here are likely to generalize well to stronger models.

The comparison layer uses the same layer function as the
composition layers (either an NN layer or an NTN layer) with
independently learned parameters and a separate nonlinearity function.
Rather than use a $\tanh$ nonlinearity here, we found better results with the leaky rectified linear function
\cite{maasrectifier}: $f(x)=\max(x, 0) +
0.01\min(x, 0)$.

To run the model forward and label a pair of phrases, the structure of
the lower layers of the network is assembled so as to match the tree
structures provided for each phrase, which is either included in the source data or given by a parser.
The word vectors are then looked up from the  vocabulary embedding matrix $V$ (one of the model parameters), and
the composition and comparison functions are used to pass information
up the tree and into the classifier at the top. For an objective
function, we use the negative log of the probability assigned to the
correct label with L2 regularization.

>>>>

For the sentence embedding portions of the network, we evaluate both TreeRNN models with
 the standard NN layer function \eqref{TreeRNN} and those with the more powerful neural tensor 
 network layer function
\eqref{TreeRNTN} proposed in \newcite{chen2013learning}. The nonlinearity $f(x) = \tanh(x)$ is applied
 elementwise to the output of either layer function.
%
\begin{gather} 
\label{TreeRNN}
\vec{y}_{\textit{TreeRNN}} = f(\mathbf{M} \colvec{2}{\vec{x}^{(l)}}{\vec{x}^{(r)}} + \vec{b}\,) \\
\label{TreeRNTN} 
\vec{y}_{\textit{TreeRNTN}} = \vec{y}_{\textit{TreeRNN}} + f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)})
\end{gather} 
%
Here, $\vec{x}^{(l)}$ and $\vec{x}^{(r)}$ are the column vector
representations for the left and right children of the node, and
$\vec{y}$ is the node's output.  The TreeRNN concatenates them, multiplies
them by an $n \times 2n$ matrix of learned weights, and adds a bias $\vec{b}$. 
The TreeRNTN adds a learned third-order tensor 
$\mathbf{T}$, of dimension $n \times n \times n$, modeling
multiplicative interactions between the child vectors. 
The comparison layer uses the same layer function as the
composition layers (either an NN layer or an NTN layer) with
independently learned parameters and a separate nonlinearity function.
Rather than use a $\tanh$ nonlinearity here, we found better results with the leaky rectified linear function
\cite{maasrectifier}: $f(x)=\max(x, 0) +
0.01\min(x, 0)$.

Other strong tree-structured networks have been proposed in past work
\cite{sochergrounded,kalchbrenner2014convolutional,irsoydeep}, but
we believe that these two provide a valuable case study, and that positive results on 
here are likely to generalize well to stronger models.

To run the model forward, we assemble the two tree networks so as to match the structures provided for each phrase, which is either included in the source data or given by a parser.
The word vectors are then looked up from the vocabulary embedding matrix $V$ (one of the model parameters), and
the composition and comparison functions are used to pass information
up the tree and into the classifier. For an objective
function, we use the negative log of the probability assigned to the
correct label with L2 regularization.

### 2015-02-27 writing/F'14 paper/quantifiers.tex
Paper update.

We have seen that recursive models can learn an approximation of propositional
logic.  However, natural languages can express functional meanings of
considerably greater complexity than this.  As a key test of whether our models 
can capture this complexity, we now directly measure the degree to which they are able to
develop suitable representations for the semantics of natural language
quantifiers like \ii{most} and \ii{all} as they interact with negation. Quantification 
and negation are far from the only place in natural language where complex functional meanings
are found, but they are natural focus, since they have
formed a standard case study in prior formal work on natural
language inference \cite{Icard:Moss:2013:LILT}..

\paragraph{Experiments}
Our experimental data consist of pairs of sentences generated
from a grammar for a simple English-like artificial language.
Each sentence contains a quantifier, a noun
which may be negated, and an intransitive verb which may be
negated. We use the basic quantifiers \ii{some}, \ii{most}, \ii{all},
\ii{two}, and \ii{three}, and their negations \ii{no}, \ii{not-all},
\ii{not-most}, \ii{less-than-two}, and \ii{less-than-three}. We also
include five nouns, four intransitive verbs, and the negation symbol
\ii{not}. In order to be able to define relations between sentences
with differing lexical items, we define the lexical relations between
each noun--noun pair, each verb--verb pair, and each
quantifier--quantifier pair. The grammar generates pairs of
sentences of this form and calculates the relations holding
between them.  Some examples of these data are provided in (\ref{p1}--\ref{p3}) below.

@@

We use the same summing baseline as in the previous section.
The highly consistent  sentence structure in this experiment means
that that model can potentially distinguish two valid sentences unless they
 differ only in the placement of the word \ii{not}, potentially introducing errors into 
the 58.8\% of examples that contain a sentence with a single
instance of \ii{not}.

@@

    \toprule
    ~ & \multicolumn{2}{c}{Train} & \multicolumn{2}{c}{Test} \\
    \midrule
    $\natind$ only &	35.4 & (7.5)	& 35.4	& (7.5)\\
    25d SumNN	&	96...&	(95...)&	90...&	(80...)\\	
    25d TreeRNN	&	\textbf{100}&	\textbf{(100)}&	99.5&	(99.3)\\
    25d TreeRNTN	&	\textbf{99.4}&\textbf{(99.5)}&	\textbf{99.4} & (99.2)\\
    \bottomrule
  \end{tabular}

  \caption{Performance on the quantifier experiments, given as \% correct and macroaveraged F1.}
  \label{qresultstable}
\end{table} 
% TODO: Fill in baseline results

%
% do not allow a blank line --- adds too much space
%
\paragraph{Results} The results, reported in Table~\ref{qresultstable}, show that both recursive models are able to learn to generalize the underlying logic satisfactorily, and that the baseline model generalizes well, but far from perfectly. We did not observe any consistent pattern in the few errors made by either model, and errors did not tend to be consistent across runs of the experiment, suggesting that there is no fundamental obstacle to learning a perfect model for this problem.

% TODO: Baseline

>>>>

We have seen that recursive models can learn an approximation of propositional
logic.  However, natural languages can express functional meanings of
considerably greater complexity than this.  As a key test of whether our models 
can capture this complexity, we now directly measure the degree to which they are able to
develop suitable representations for the semantics of natural language
quantifiers like \ii{most} and \ii{all} as they interact with negation. Quantification 
and negation are far from the only place in natural language where complex functional meanings
are found, but they are natural focus, since they have
formed a standard case study in prior formal work on natural
language inference \cite{Icard:Moss:2013:LILT}.

\paragraph{Experiments}
Our data consist of pairs of sentences generated
from a grammar for a simple English-like artificial language.
Each sentence contains a quantifier, a noun
which may be negated, and an intransitive verb which may be
negated. We use the quantifiers \ii{some}, \ii{most}, \ii{all},
\ii{two}, and \ii{three}, and their negations \ii{no}, \ii{not-all},
\ii{not-most}, \ii{less-than-two}, and \ii{less-than-three}, and also
include five nouns, four intransitive verbs, and the negation symbol
\ii{not}. In order to be able to define relations between sentences
with differing lexical items, we define the lexical relations for
each noun--noun pair, each verb--verb pair, and each
quantifier--quantifier pair. The grammar then generates pairs of
sentences and calculates the relations
between them.  
Some examples of these data are provided in (\ref{p1}--\ref{p3}) below.

@@

We use the same summing baseline as in \S\ref{sec:recursion}.
The highly consistent  sentence structure in this experiment means
that that model can potentially distinguish two valid sentences unless they
 differ only in the placement of \ii{not}, potentially introducing errors on 
the 58.8\% of examples that contain a sentence with a single
instance of \ii{not}.

@@

    \toprule
    ~ & \multicolumn{2}{c}{Train} & \multicolumn{2}{c}{Test} \\
    \midrule
    $\natind$ only &	35.4 & (7.5)	& 35.4	& (7.5)\\
    25d SumNN	&	97.0&	(97.2)&	89.6&	(91.1)\\	
    25d TreeRNN	&	\textbf{100}&	\textbf{(100)}&	\textbf{99.4}&	\textbf{(99.5)}\\
    25d TreeRNTN	&	99.4&(99.5)&	99.4 & (99.2)\\
    \bottomrule
  \end{tabular}

  \caption{Performance on the quantifier experiments, given as \% correct and macroaveraged F1.}
  \label{qresultstable}
\end{table}

%
% do not allow a blank line --- adds too much space
%
\paragraph{Results} The results (Table~\ref{qresultstable}) show that both tree models are able to learn to generalize the underlying logic almost perfectly. The baseline summing model can largely memorize the training data, but does not generalize well. We do not find any consistent pattern in the handful of errors made by either tree model, and no errors were consistent across model restarts, suggesting that there is no fundamental obstacle to learning a perfect model for this problem.

### 2015-02-27 writing/F'14 paper/recursion.tex
Paper update.

\section{Recursive structure}\label{sec:recursion}
% TODO: Enter baseline results, use [[Tree]]RNN.

% TODO: Clean up next sent.
Recursive structure is a prominent feature of natural language. 
This section examines the degree to which our models are able to learn a 
compositional semantics over recursive structures.
In our evaluations, we exploit the fact that our logical language
is infinite by testing on strings that are longer and more complex
than any seen in training.

@@

\paragraph{Experiments}
As in \S\ref{sec:join}, we test this phenomenon on data generated from a logic,
 but we now replace the unanalyzed symbols
from that experiment with complex formulae. These formulae
represent a truth-functionally complete classical propositional logic:
each atomic symbol is a variable over the domain \{$\True$, $\False$\}, and the only
operators are truth-functional ones.  Table~\ref{tab:pl} defines this
logic, and Table~\ref{tab:plexs} gives examples of the some of the shorter relational
statements that we generate from it as train and test data. To compute these relations
between statements, we exhaustively enumerate the sets of assignments
of truth values to propositional variables that would satisfy each of
the statements, and then we convert the set-theoretic relation between
those assignments into one of the seven relations in
Table~\ref{b-table}. As a result, each relational statement represents
a valid theorem of the propositional logic, and and to succeed, the models 
must learn to reproduce the behavior of a theorem prover.\footnote{
Socher et al.~\shortcite{socher2012semantic} show that a matrix-vector TreeRNN
model somewhat similar to our TreeRNTN can learn boolean logic, 
a logic where the atomic symbols are simply the
values $\True$ and $\False$. While learning the operators of that logic is not trivial, the outputs of
each operator can be represented accurately by a single bit.
In the more demanding task presented here, the atomic symbols are variables over these values, and the sentence vectors must thus be able to distinguish up to $2^{64}$ distinct conditions on valuations.}

@@

In our experiments, we randomly generate unique pairs 
of formulae and compute the relation that holds for each pair.
We discard pairs in which either statement is either a tautology or a
contradiction, for which the seven relations in
Table~\ref{b-table} are undefined. The resulting set of formula pairs is
then partitioned into 13 bins according the number of logical operators in
the longer of the two formulae. We then sample 20\% of each
bin for a held-out test set.

\begin{figure*}[t]
  \centering
  \includegraphics[width=5.75in]{decayfig.eps}
%  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
  \caption{Model performance on propositional logic, by expression size.}
%    The dashed line indicates that only expressions of size four or less appeared in the training data. \textbf{Bottom:} Semantically distinct formulae should have different
%    representations. As measured by Euclidean distance, only the TreeRNTN
%    achieves this for formulae containing more than a small number of
%    connectives (\ii{and} in this example). The TreeRNN quickly collapses the representations,
%    failing to capture the meaning contrasts.

  \label{prop-results}

\end{figure*}

If we do not implement any constraint that the two statements being
compared are similar in any way, then the generated data are dominated
by statements in which the two formulae refer to largely separate
subsets of the six variables, which means that the $\natind$ relation
is almost always correct.  In an effort to balance the distribution of
relation labels without departing from the basic task of modeling
propositional logic, we disallow individual pairs of statements from
referring to more than four of the six propositional variables.

In order to test the model's generalization to unseen structures, we discard
training examples with more than 4 connectives, yielding 60k short training examples,
and 21k test examples containing up to 12 connectives.
We evaluated both the TreeRNN and TreeRNTN on the resulting data, initializing the model parameters randomly,
including the vector representations of the six variables. 
As a baseline, we also train a summing neural network
which is largely identical to the TreeRNN, except that instead of learning a parametric composition function,
it simply sums the term vectors in each expression to compose them. Unlike the two models under study,
whose ability to accurately simulate the logic is an open question, this baseline does not use word order,
and is as such guaranteed to ignore some information that it would need to succeed perfectly.

% TODO: Insert baseline result.
\paragraph{Results} Fig.~\ref{prop-results} shows the relationship
between test accuracy and statement length. BASELINE BASELINE
 BASELINE BASELINE BASELINE BASELINE BASELINE BASELINE BASELINE   We found that both 
models were able to perform well on unseen small test examples, 
with TreeRNN accuracy above
98\% and TreeRNTN accuracy above 99\% on formulae below length five.
Starting at size four, performance gradually falls with increasing
size. Training accuracy was 99.4\% for the TreeRNN and 99.8\% for the TreeRNTN.

The performance of these models on small unseen test examples
indicates that they learned correct approximations of the underlying
logic. These approximations are accurate enough to yield
correct answers when the composition layer is only applied a small
number of times, but that the error in the approximation grows with
increasing expression size, resulting in gradually dropping performance.

It appears that the TreeRNTN model overfit somewhat relative to the plain TreeRNN,
performing better on the training data and short examples, but worse on longer
examples. Two factors are likely involved in this: even with the lower dimension,
the TreeRNTN composition function has about eight times as many parameters as the
TreeRNN, and thus more capacity to find a solution involving a large number of narrow 
generalizations that work on the training data but don't generalize well. In addition,
the TreeRNTN used here has weaker L2 regularization than the 
TreeRNN ($\lambda = 0.0003$ vs. $0.001$). This
yielded better performance averaged across all bins, but may have facilitated this effect.

Though imperfect, model performance on the most difficult set of expressions is nonetheless striking. 
The weaker TreeRNTN classifies true examples of every class but $\nateq$ 
(which is rare in long examples, and occurs only once here) correctly the majority of the time.
% TODO: Better present examples.

>>>>

\section{Recursive structure}\label{sec:recursion}

Recursive structure is a prominent feature of natural language. 
This section examines the degree to which our models can learn a 
compositional semantics over recursive structures.
In our evaluations, we exploit the fact that our target language
is infinite by testing on strings that are longer and more complex
than any seen in training.

@@

\paragraph{Experiments}
As in \S\ref{sec:join}, we generate artificial data from a formal system,
 but we now replace the unanalyzed symbols
from that experiment with complex formulae. These formulae
represent a complete classical propositional logic:
each atomic symbol is a variable over the domain \{$\True$, $\False$\}, and the only
operators are truth-functional ones.  Table~\ref{tab:pl} defines this
logic, and Table~\ref{tab:plexs} gives some short examples relational statements from our data.
 To compute these relations
between statements, we exhaustively enumerate the sets of assignments
of truth values to propositional variables that would satisfy each of
the statements, and then we convert the set-theoretic relation between
those assignments into one of the seven relations in
Table~\ref{b-table}. As a result, each relational statement represents
a valid theorem of the propositional logic, and to succeed, the models 
must learn to reproduce the behavior of a theorem prover.\footnote{
Socher et al.~\shortcite{socher2012semantic} show that a matrix-vector TreeRNN
model somewhat similar to our TreeRNTN can learn boolean logic, 
a logic where the atomic symbols are simply the
values $\True$ and $\False$. While learning the operators of that logic is not trivial, the outputs of
each operator can be represented accurately by a single bit.
In the more demanding task presented here, the atomic symbols are variables over these values, and the sentence vectors must thus be able to distinguish up to $2^{64}$ distinct conditions on valuations.}

@@

In our experiments, we randomly generate unique pairs 
of formulae and compute the relation that holds for each pair.
We discard pairs in which either statement is either a tautology or a
contradiction, for which the seven relations in
Table~\ref{b-table} are undefined. The resulting set of formula pairs is
then partitioned into 13 bins according the number of logical operators in
the longer of the two formulae. We then sample 20\% of each
bin for a held-out test set.
If we do not implement any constraint that the two statements being
compared are similar in any way, then the generated data are dominated
by statements in which the two formulae refer to largely separate
subsets of the six variables, which means that the $\natind$ relation
is almost always correct.  In an effort to balance the distribution of
relation labels without departing from the basic task of modeling
propositional logic, we disallow individual pairs of statements from
referring to more than four of the six propositional variables.

In order to test the model's generalization to unseen structures, we discard
training examples with more than 4 logical operators, yielding 60k short training examples,
and 21k test examples containing up to 12 operators.
In addition to the two tree models, we also train a summing NN baseline
which is largely identical to the TreeRNN, except that instead of using a learned composition function,
it simply sums the term vectors in each expression to compose them. Unlike the two tree models, this baseline does not use word order,
and is as such guaranteed to ignore some information that it would need to succeed perfectly.

\begin{figure}[t]
  \centering
  \includegraphics[width=3.05in]{decayfig.pdf}
%  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
  \caption{Results on recursive structure. The ``$\natind$ only'' line shows the frequency of the most frequent class, $\natind$, and the vertical dotted line marks the size of the longest training examples.}
%    The dashed line indicates that only expressions of size four or less appeared in the training data. \textbf{Bottom:} Semantically distinct formulae should have different
%    representations. As measured by Euclidean distance, only the TreeRNTN
%    achieves this for formulae containing more than a small number of
%    connectives (\ii{and} in this example). The TreeRNN quickly collapses the representations,
%    failing to capture the meaning contrasts.

  \label{prop-results} 
\end{figure}

\paragraph{Results} Fig.~\ref{prop-results} shows the relationship
between test accuracy and statement length. While the summing baseline model performed poorly across the board, we found that both recursive
models were able to perform well on unseen small test examples, 
with TreeRNN accuracy above
98\% and TreeRNTN accuracy above 99\% on formulae below length five, indicating
that they learned correct approximations of the underlying
logic. Training accuracy was 66.6\% for the SumNN, 99.4\% for the TreeRNN, and 99.8\% for the TreeRNTN.

After the size four training cutoff, performance gradually decays with expression size for both tree models, suggesting that the learned approximations were accurate but lossy.
Despite its stronger performance on short sentences, but the TreeRNTN's performance
decayed more quickly than the TreeRNN's. 
This suggests to us that it learned to interpret many specific fixed-size tree structures directly,
allowing it to get away without learning strong generalizations about how to compose
individual terms in a more general way.
Two factors may have contributed to the learning of these narrower generalizations: 
even with the lower dimension,
the TreeRNTN composition function has about eight times as many parameters as the
TreeRNN, and the TreeRNTN worked best with weaker L2 regularization than the 
TreeRNN ($\lambda = 0.0003$ vs. $0.001$). Nonetheless, the performance of both 
models on even the longest set of test sentences is good enough to indicate that
 both can still perform at least least a rough approximation of the underlying theorem proving task.
